---
title: "Motor Trend Analysis"
author: "Andrey Alferov"
date: "Sunday, May 24, 2015"
output:
  pdf_document:
    latex_engine: xelatex
---

## Executive summary
You work for Motor Trend, a magazine about the automobile industry. Looking at a data set of a collection of cars, they are interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome). They are particularly interested in the following two questions:

1. “Is an automatic or manual transmission better for MPG”
2. “Quantify the MPG difference between automatic and manual transmissions”

## Data Processing
The data was extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973–74 models).

```{r, warning=FALSE}
library(datasets)
library(ggplot2)
library(dplyr)

data(mtcars) #Load mtcars dataset
```

Let's have a brief look at mtcars dataset.
```{r}
head(mtcars)

dim(mtcars)

summary(mtcars)
```

The dataset consists of 32 observations on 11 variables. 

## Analysis

### Is an automatic or manual transmission better for MPG?
To answer that question lets compare average mpg for automatic and manual transmissions.
```{r}
mean(filter(mtcars, am == 1)$mpg) #Calculate mean mpg for manual transmission
mean(filter(mtcars, am == 0)$mpg) #Calculate mean mpg for automatic transmission
```
As we can see cars with manual transmission have greater mpg than cars with manual transmission.

Lets build a boxplot displaying mpg per transmission type.
```{r}
p <- ggplot(mtcars, aes(factor(am), mpg))
p + geom_boxplot() + geom_jitter()
```
The plot also prooves that cars with manual transmission type (represented by 1) has a higher mean for mpg than automatic (represented by 0).

Lets perform a t-test to confirm the null hypothesis that transmission type affects mpg.
```{r}
t.test(mtcars$mpg ~ mtcars$am, conf.level=0.95)
```

p-value = 0.001374 is less then 0.05, so we reject the null hypothesis that there is no difference in MPG per transmission type.

### Quantify the MPG difference between automatic and manual transmissions.

Lets check how all variables affect mpg difference by fitting corresponding linear model.
```{r}
fit_all = lm(data = mtcars, mpg ~ .)
summary(fit_all)
```
As we can see there are 2 predictor variables: wt (Weight) and am (Transmission type) which affects mpg variance at most.

Lets try to choose the best set of predictors automatically.
```{r}
fit_best = step(lm(data = mtcars, mpg ~ .), trace=0)
summary(fit_best)
```
It looks like the best model is the one that includes wt, qsec and am, which means besides transmission types, weight and accelearation also needs to be considered. Weight negatively changes with mpg, and qsec and am positively changes. Every lb/1000 weight increase will cause a decrease of roughly 4 mpg, every increase of 1/4 mile time will cause an increase of 1.2 mpg, and on average, manual transmission is 2.9 mpg better than automatic transmission. The model is able to explain 85% of variance. The residual plots also seems to be randomly scattered (see appendix).

###Conclusion
As a result of analysis it can be concluded that manual transmission is better than automatic transmission by 2.9 mpg. However, transmission type is not the only factor accounting for MPG, weight, and acceleration (1/4 mile time) also needs to be considered.

###Appendix

Residual plots

```{r}
plot(fit_best)
```
