---
title: "Motor Trend Analysis"
author: "Andrey Alferov"
date: "Sunday, May 24, 2015"
output:
  pdf_document:
    latex_engine: xelatex
---

## Executive summary
You work for Motor Trend, a magazine about the automobile industry. Looking at a data set of a collection of cars, they are interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome). They are particularly interested in the following two questions:

1. “Is an automatic or manual transmission better for MPG”
2. “Quantify the MPG difference between automatic and manual transmissions”

## Data Processing
The data was extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973–74 models).

```{r, warning=FALSE}
library(datasets)
library(ggplot2)
library(dplyr)

data(mtcars) #Load mtcars dataset
```

Let's have a brief look at mtcars dataset.
```{r}
head(mtcars)

dim(mtcars)

summary(mtcars)
```

The dataset consists of 32 observations on 11 variables. 

## Analysis

### Is an automatic or manual transmission better for MPG?
To answer that question lets compare average mpg for automatic and manual transmissions.
```{r}
mean(filter(mtcars, am == 1)$mpg) #Calculate mean mpg for manual transmission
mean(filter(mtcars, am == 0)$mpg) #Calculate mean mpg for automatic transmission
```
As we can see cars with manual transmission have greater mpg than cars with manual transmission.

Lets build a boxplot displaying mpg per transmission type.
```{r}
p <- ggplot(mtcars, aes(factor(am), mpg))
p + geom_boxplot() + geom_jitter()
```
The plot also prooves that cars with manual transmission type (represented by 1) has a higher mean for mpg than automatic (represented by 0).

Lets perform a t-test to confirm the null hypothesis that transmission type affects mpg.
```{r}
t.test(mtcars$mpg ~ mtcars$am, conf.level=0.95)
```

p-value = 0.001374 is less then 0.05, so we reject the null hypothesis that there is no difference in MPG per transmission type.

### Quantify the MPG difference between automatic and manual transmissions.


